---
title: "NB03_EDA_Continued"
output: html_document
---

#===Preamble===#


```{r}
# Clear environment of variables
#rm(list = ls())
``` 

## Import Packages
```{r}
# Directory Access
library(rstudioapi) # to change the working directory respectively

# Data Manipulation
library(jsonlite)# for reading in json data and credentials to the purpose-built database
library(dplyr) # to clean data

# Plotting and Visualisation
library(ggplot2) # to plot nice histograms
library(grid) # for more plotting interaction
library(zoo) # for time series plotting
library(patchwork) # for combining figures
library(xtable) # to write out tables into latex
library(viridisLite)  # For viridis color palette


# Database Access
library(DBI)
library(RPostgres)

```



## Set up the right path
```{r}
# Get the path of the current script
current_file_path <- rstudioapi::getActiveDocumentContext()$path
# Extract the directory from the file path
current_file_dir <- dirname(current_file_path)
# Set the current working directory to the directory of the current file
setwd(current_file_dir)
```

## Database

###  Get the credentials which allow us access to the database

```{r}
# Build the path to the JSON credentials file (1 level up)
credentials_file_path <- file.path(current_file_dir, '..', 'supabase_credentials.json')

# Read the JSON file
credentials <- fromJSON(credentials_file_path)
```



### Call in the source file that allows access to the database

```{r}
# Call in the source file that allows access to the database
source("../src/macro_utils/sql_queries.R")
```




```{r}
# Establish connection to the database
con <- get_supabase_connection(
  user = "postgres",
  password = credentials$password,
  host = credentials$host,
  port = 5432,
  database = "postgres"
)


```

### Load the Data

```{r}
# Extract the data of interest
plot_data <- dbGetQuery(con, "SELECT * FROM growth_data")
```

```{r}
plot_data
```

### Post Database Extraction Data Cleaning

```{r}
# Convert the date to date format
plot_data$date <- as.Date(plot_data$date)


# Sort the data by country and date
plot_data <- plot_data %>%
  arrange(country, date)

# Add GFC and COVID indicator columns
plot_data <- plot_data %>%
  mutate(
    gfc = date >= as.Date("2008-06-01") & date <= as.Date("2009-12-31"),
    covid = date >= as.Date("2019-12-01") & date <= as.Date("2021-12-31")
  )
```

### Visualise the different time series for different groupings (countries)


#### Additional Data Cleaning for Plotting Purposes Only

```{r}
# Convert GDP (PPP) to billions for better readability
plot_data <- plot_data %>%
  mutate(
    gdp_ppp = gdp_ppp / 1e9,       # GDP in billions
    labour_force = labour_force / 1e6  # Labour force in millions
  )


```

```{r}
# Calculate min and max date for each aggregate shock period (GFC and COVID)
# Create a named list of period labels and their associated logical columns
periods <- list(GFC = "gfc", COVID = "covid")

# Create a data frame of shock periods
shock_ranges <- bind_rows(lapply(names(periods), function(label) {
  indicator <- periods[[label]]
  data_subset <- plot_data %>% filter(.data[[indicator]])
  
  if (nrow(data_subset) > 0) {
    tibble(
      period = label,
      xmin = min(data_subset$date, na.rm = TRUE),
      xmax = max(data_subset$date, na.rm = TRUE)
    )
  } else {
    tibble(
      period = label,
      xmin = as.Date(NA),
      xmax = as.Date(NA)
    )
  }
}))

# Create A colour mapping for these shock periods
shock_fill_mapping <- c(
  GFC = "#0457db",    # orange
  COVID = "#87CEEB"   # light blue
)
```
```{r}
shock_ranges
```

#### Setup the Plotting Environment (Colour Mappings, X axis breaks etc)

```{r}
# 1. Get unique countries
countries <- unique(plot_data$country)
num_countries <- length(countries)

# 2. Generate a color palette using viridis and convert to HEX
# viridisLite::viridis returns HEX by default
color_palette <- viridisLite::viridis(num_countries)

# 3. Map colors to countries
# Create a named vector where names are country names and values are HEX colors
color_mapping <- setNames(color_palette, countries)
```
```{r, eval=F}
color_mapping
```


```{r}
# 1. Select relevant columns and remove duplicates
label_df <- plot_data %>%
  select(date, year) %>%
  distinct() %>%
  arrange(date)

# 2. Choose every 2nd row (adjust the step size as needed)
label_df_filtered <- label_df[seq(1, nrow(label_df), by = 2), ]

# 3. Extract breaks and labels
yq_breaks <- label_df_filtered$date
yq_labels <- label_df_filtered$year
```

### Plot Key Data Series

```{r}
# Filter only the countries of interest for illustrative purposes
plot_data <- plot_data %>%
  filter(country %in% c("United States", "Germany", "Japan"))
```

```{r}
# Get range of each variable
range_primary <- range(plot_data$gdp_ppp, na.rm = TRUE)
range_secondary <- range(plot_data$labour_force, na.rm = TRUE)

# Rescale labour_force to match the GDP scale
plot_data <- plot_data %>%
  mutate(labour_force_rescaled = (labour_force - range_secondary[1]) / diff(range_secondary) * diff(range_primary) + range_primary[1])
```


```{r}
# Create the plot
g <- ggplot(plot_data, aes(x = date)) +
  
  # GDP Lines per country
  geom_line(aes(y = gdp_ppp, color = country), linewidth = 1) +

  # Labour Force Lines (dashed) per country
  geom_line(aes(y = labour_force_rescaled, color = country), linetype = "dashed", linewidth = 1) +
  
  # Add shaded rectangles for shock periods
  geom_rect(
    data = shock_ranges,
    inherit.aes = FALSE,
    aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = period),
    alpha = 0.15
  ) +
  
  # SUPPORT 2 AXES
  # Primary y-axis
  scale_y_continuous(
    name = "GDP (PPP, USD Billions)",

    # Secondary axis transformation
    sec.axis = sec_axis(
        transform = ~ (.-range_primary[1]) / diff(range_primary) * diff(range_secondary) + range_secondary[1],
        name = "Labour Force (Millions)"
      )
  ) +
  
  # Use same color mapping
  scale_color_manual(name = "Country", values = color_mapping) +
  
  # Manually assign fill colors for shock periods
  scale_fill_manual(name = "Shock Period", values = shock_fill_mapping) +
  
  # Custom x-axis breaks and labels
  scale_x_date(breaks = yq_breaks, labels = yq_labels) +
  
  # Add titles and axis labels
  labs(
    title = "Key Data Comparison Across Countries",
    x = "Year",
    color = "Country"
  ) +
  
  # Add white Background
  theme_minimal() +
  # Rotate x-axis text
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.5)
  ) +
  
  # Remove linetype guide
  guides(linetype = "none")

print(g)
```

```{r}
plot_dual_axis_comparison <- function(
  df,
  countries,
  primary_var,
  secondary_var,
  primary_label = "Primary Variable",
  secondary_label = "Secondary Variable",
  y_title = "Year",
  two_axes = T
) {
  # -------------------------
  # Filter for selected countries
  # -------------------------
  df <- df %>%
    filter(country %in% countries)

  # -------------------------
  # Capture data ranges
  # -------------------------
  range_primary <- range(df[[primary_var]], na.rm = TRUE)
  range_secondary <- range(df[[secondary_var]], na.rm = TRUE)

  # -------------------------
  # Rescale secondary variable to match primary axis
  # -------------------------
  df <- df %>%
    mutate(
      secondary_rescaled = (.[[secondary_var]] - range_secondary[1]) / diff(range_secondary) * 
                           diff(range_primary) + range_primary[1]
    )

  # -------------------------
  # Build long-format data with `measure` column for linetype mapping
  # -------------------------
  df_primary <- df %>%
    mutate(value = .[[primary_var]], measure = primary_label)

  df_secondary <- df %>%
    mutate(value = secondary_rescaled, measure = secondary_label)

  df_long <- bind_rows(df_primary, df_secondary)
  
  # Create the y-axis scale depending on whether two_axes is TRUE
  y_axis_scale <- if (two_axes) {
    scale_y_continuous(
      name = primary_label,
      sec.axis = sec_axis(
        trans = ~ (.-range_primary[1]) / diff(range_primary) * diff(range_secondary) + range_secondary[1],
        name = secondary_label
      )
    )
  } else {
    scale_y_continuous(name = primary_label)
  }

  # -------------------------
  # Generate color mapping per country using viridis palette
  # -------------------------
  country_list <- unique(df$country)
  num_countries <- length(country_list)
  color_palette <- scales::viridis_pal()(num_countries)
  color_mapping <- setNames(color_palette, country_list)
  
  # Create named vector for linetype mapping
  linetype_mapping <- setNames(c("solid", "dashed"), c(primary_label, secondary_label))


  # -------------------------
  # Build the ggplot
  # -------------------------
  g <- ggplot(df_long, aes(x = date, y = value, color = country, linetype = measure)) +

    # Line plot for both measures
    geom_line(linewidth = 1) +

    # Optional: Add shaded rectangles for known shock periods
    geom_rect(
      data = shock_ranges,
      inherit.aes = FALSE,
      aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = period),
      alpha = 0.15
    ) +

    # Y-axis scale (optionally dual)
    y_axis_scale +
    

    # Apply color and fill mappings
    scale_color_manual(name = "Country", values = color_mapping) +
    scale_fill_manual(name = "Shock Period", values = shock_fill_mapping) +

    # Manual legend for linetype
    scale_linetype_manual(
    name = "Measure",
    values = linetype_mapping
    ) +

    # Custom x-axis ticks
    scale_x_date(breaks = yq_breaks, labels = yq_labels) +

    # Labels and theme
    labs(
      title = paste(primary_label, "and", secondary_label, "Comparison"),
      x = y_title
    ) +

    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 0.5)
    )

  return(g)
}

```
```{r, fig.width=16, fig.height = 6}
g<- plot_dual_axis_comparison(
  df = plot_data,
  countries = c("United States", "Germany", "Japan"),
  primary_var = "gdp_ppp",
  secondary_var = "labour_force",
  primary_label = "GDP (PPP, USD Billions)",
  secondary_label = "Labour Force (Millions)",
  two_axes = T
)
print(g)
```



